\documentclass[a4paper]{article}
    \input{../notesheader.tex}
    \usepackage[normalem]{ulem}

    \newcommand{\chapternumber}{12}
    \newcommand{\chaptertitle}{Antebellum Culture and Reform}

    \input{../titleheader.tex}
    \begin{document}
    \maketitle
    \SetBgContents{\rule[0em]{4pt}{\textheight}}
    \cornell[Key Concepts]{What are this chapter's key concepts?}{\begin{itemize}
        \item \textbf{4.1.II.B} - A culture formed blending national American and European elements with regional sections
        \item \textbf{4.1.II.C} - Romantic and liberal social beliefs influenced literature, art, philosophy, and architecture
        \item \textbf{4.1.II.D} - Enslaved/free Afr. Americans $\to$ communities to protect dignity, joining pol. efforts to change status
        \item \textbf{4.1.III.A} - Americans created organizations to improve society thru. reform
        \item \textbf{4.1.III.C} - Women's rights movement developed for gender equality, culminating in Seneca Falls Convention
        \item \textbf{4.3.II.B} - North saw $\uparrow$ antislavery while South (despite few owning slaves) saw $\uparrow$ slavery as natural way of life
    \end{itemize}}
    \cornell[The Romantic Impulse]{How did a unique culture develop in American society?}{\textbf{Painting reflected nationalistic ideals through the power of wild natural environments, Northern literature reflecting ideals of independence, liberty, and democracy, Southern literature based either around the wealthy aristocrats or those inhabiting the fringes of society as rural peasants, and the Transcendentalists, focusing on an appreciation of the surrounding world from a personal standpoint, came together to form a unique American literary style. Nature became the focal point of a large part of society, a belief which often manifested itself in utopian societies where inhabitants worked in harmony with their natural environment. These societies, and others, often distorted gender roles by going against social norms to give women significantly more, or different forms of, power. Mormonism attracted those who felt out of touch or increasingly distant from material society, and sought to create an idealistic society based around human perfection.}}
    \cornell{How did American painting reflect nationalism and romanticism?}{\begin{itemize}
        \item Sydney Smith, English wit, expressed that no one outside of America enjoyed American art; but U.S. enjoyed significantly
        \item Most popular aimed to show landscapes: not mere countryside, but instead wildest places (w/ "sublime": awe/fear of nature)
        \begin{itemize}
            \item Frederic Church, Thomas Cole, Thomas Doughty, Asher Durand, all of NY, known as Hudson River School, painted rugged Hudson Valley
            \begin{itemize}
                \item Felt nature greatest source of wisdom; stre
            \end{itemize}
            \item Many began to travel westward to witness spectacular world of Yosemite Valley, Yellowstone, Rockey Mountains
            \begin{itemize}
                \item Thomas Moran, Albert Bierstadt traveled throughout country
            \end{itemize}
        \end{itemize}
    \end{itemize}
    \textbf{American painting, despite not having reached an international audience, appealed greatly to Americans themselves. It generally focused on the idea of "wild nature" and the sublime, initially centered around the Hudson Valley in NY but extending westward.}}
    \cornell{How did American writers generally emphasize ideas of liberty?}{\begin{itemize}
        \item GB's Sir Walter Scott most popular in early nineteenth century; most common American novels were "sentimental novels" of women
        \item James Fenimore Cooper stressed ideals of wilderness, adventure, growing up on frontier NY ("Leatherstocking Tales")
        \begin{itemize}
            \item Represented ideal for true American literature, also depicting central social concerns like fear of disorder, ideal of independence
        \end{itemize}
        \item Walt Whitman, "poet of American democracy," born in 1819 w/ start as newspaper apprentice
        \begin{itemize}
            \item Founded and led NY newspaper \textit{Long Islander}
            \item Printed first volume of work \textit{Leaves of Grass} in 1855, celebrating democracy/individualism; reflected homosexuality in intolerant society
        \end{itemize}
        \item Herman Melville sailed world before rooting back in U.S. and publishing \textit{Moby Dick}, portraying Ahab, captain of whaling vessel, seeking violent whale Moby Dick for fulfillment
        \begin{itemize}
            \item Spirit ultimately -> annihilation
        \end{itemize}
        \item Edgar Allen Poe, one of few southern writers, created sad stories, with books and famous poem "The Raven," seeking to transcend from intellect, explore emotion; had great effect on other poets
    \end{itemize}
    \textbf{Although British writer Walter Scott was popular, American writers like Cooper, describing the independence of the wilderness, Whitman, stressing individualism and democracy through his poems reflecting his troubled state as a homosexual man in an intolerant society, Melville, whose \textit{Moby Dick} revealed the potential destructive nature of the human spirit, and Poe's sad poems exploring true emotion beyond intellect gradually grew in popularity.}}
    \cornell{What were the critical ideals of literature in the South?}{
        Antebellum Southern literature was based around defining the American nation, but often contradicted the true state of society.
        \begin{itemize}
            \item Novelists created romances/eulogies describing upper South plantation system 
            \begin{itemize}
                \item Early (1830s) from Richmond, including Beverly Tucker, William Alexander Caruthers, John Pendleton Kennedy 
                \item Literary capital moved to Charleston in 1840s w/ William Gilmore Simms expressing nationalism initially hoping to transcend regional diffs. but soon defended slavery 
            \end{itemize}
            \item Writers on fringes of plantation society depicted backwoods societies
            \begin{itemize}
                \item Included Augustus B. Longstreet, Joseph G. Baldwin, Johnson J. Hooper
                \item Centered around ordinary, poor people with unique humor; Mark Twain most powerful of group
            \end{itemize}
        \end{itemize}
        \textbf{Southern literature aimed to define the American nation: novelists generally focused on the cavaliers, initially mostly from Richmond but shifting to Charleston. Another group of fringe writers described the impoverished instead of the aristocratic class.}}
        \cornell{Who were the Transcendentalists?}{\begin{itemize}
            \item Transcendentalists focused on individualism by distinguishing "reason" (innate ability of all to understand beauty/truth with full expression of emotions) and "understanding" (intellect applied to narrow confines of society)
            \item Leader was Emerson, lecturer devoted to sharing beliefs, speaking with intellectuals daily 
            \begin{itemize}
                \item Produced some poetry but known for essays/lectures like "Nature" and "Self-Reliance"
                \item Nationalist: believed in cultural independence - lecture "American Scholar" argued that European cultural heritage be ignored and instictive genius be harnessed
            \end{itemize}
            \item Henry David Thoreau significant, too, arguing repression of society $\to$ desperation w/ no one conforming to social pressures
            \begin{itemize}
                \item Went to Walden Pond in Concord Woods and lived in cabin to live deliberately and simply
                \item Resisted slave-allowing govt. by not paying poll tax -> jailed briefly in 1846
                \begin{itemize}
                    \item Argued in "Resistance to Civil Government" that morality > legal codes
                \end{itemize}
            \end{itemize}
        \end{itemize}
        \textbf{The Transcendentalists argued for individualism through an innate personal inderstanding of the world over absorption of mere knowledge and its application to narrow fields. Led by Emerson, a lecturer producing powerful essays like "Nature and Self-Reliance" and a nationalist believing in cultural independence and also supported by Henry David Thoreau, who isolated himself in nature and resisted the government's allowance of slavery, the Transcendentalists produced a powerful repretoire of literature.}}
        \cornell{How did the Transcendentalists defend critical concepts of nature?}{\textbf{The Transcendentalists and others felt that nature was not a scientific virtue or an economic stimulus, but instead a place for spirituality and inspiration forming a basic part of humanity. Their work marked the beginning of the environmental movement.}}
        \cornell{How did several utopian societies emerge?}{\begin{itemize}
            \item Brook Farm: Boston Transcend. George Ripley created equal social organization
            \begin{itemize}
                \item One of first thinkers to establish leisure as beneficial, restorative practice
                \item Manual labor -> gap slowly bridged betw. nature and instinct
                \item Realistically, tensions began to rise and large fire finally split up group for good
                \item Writher N. Hawthorne one of original inhabitants, wrote \textit{The Blithesdale Romance} to describe terrrible consequences to most devout
            \end{itemize}
            \item Charles Fourier, French philosopher detailing socialist communities known as "phalanxes," inspired numerous communities
            \item Philanthropist Robert Owen founded New Harmony in Indiana devoted to cooperation; despite economic failure, continued to inspire
        \end{itemize}
        \textbf{Brook Farm, one of the earliest utopian experiments, placed numerous inhabitants together on a farm, required to perform manual labor to appreciate their natural surroundings. Fourier, a French philosopher, inspired multiple communities and Owen, a Scottish philanthropist, created the New Harmony reservation in Indiana. Although nearly all utopian societies failed economically, they continued to inspire countless.}}
        \cornell{How did changing social philosophies transform gender roles?}{\begin{itemize}
            \item Margaret Fuller, Transcendentalist close to Emerson, shunned domestic female stereotype and encouraged intellectual power and social dominance
            \item Oneida Community by John Humphrey Noys rejected traditional family ideals with everyone married to each other
            \begin{itemize}
                \item Sexual activity monitored to prevent rape; children generally raised by multiple parents
                \item Believed in ability to limit male desire by removing family bonds
            \end{itemize}
            \item Shakers, by "Mother" Ann Lee in 1770s, received their peak population in the antebellum period; known for commitment to abstinence
            \begin{itemize}
                \item Never passed onto children; always voluntary choice
                \item Men and women generally segregated but regarded as equal with gender-ambiguous God
                \item Women had majority of power
                \item Primary goal not for equitable gender roles but for society distinct from chaos of normal life
            \end{itemize}
            \item Amana Community by 1843 German immigrants began in Iowa
        \end{itemize}
        \textbf{Fuller, a Transcendentalist; the Oneida Community, replacing the notion of marriage with one of universal union; and the Shakers, requiring complete abstinence all pushed for increased women's rights to varying degrees.}}
        \cornell{Who were the Mormons?}{\begin{itemize}
            \item Mormons (Church of Jesus Christ of Latter Day Saints) aimed to create new ordered society; led by young Joseph Smith with Book of Mormon
            \begin{itemize}
                \item Smith claimed ancient prophet had written words on golden tablets shown to him by God; he had simply translated 
                \item Book argued that group of Israelites had formed a fruitful sociey in America; Jesus resurrected there
                \item After early Americans began to stray from righteous beliefs, punished by God with darkened skin, cleared memory as natives
            \end{itemize}
            \item Smith established sizable following by 1831, but continually persecuted for radical religious beliefs and goal to find isolated community
            \begin{itemize}
                \item Polygamy, social rigidity, and secrecy damaged reputation
                \item Sought human perfection achieved through social organization based around a tight hierarchical structure 
            \end{itemize}
            \item Settled in Nauvoo, IL, but Smith arrested in 1844 and imprisoned and killed by angry mob
            \item Smith's successor, Brigham Young, led society of 12,000 ppl. across desert to Salt Lake City 
            \begin{itemize}
                \item Majority of converts those who were displaced in modern society/felt lacking control in material world
            \end{itemize}
        \end{itemize}
        \textbf{Mormonism began with Joseph Smith, who claimed to have been shown works detailing the arrival of a group of Israelites to America and the formation of a long-standing, fruitful society. Hoping to recreate this, Smith created a rigid social order based around a polygamous society hoping to reach human perfection. Although Smith was imprisoned and killed during his first true attempt at settlement, his successor, Brigham Young, led 12,000 to Salt Lake City.}}
        \cornell[Remaking Society]{What major reform movements emerged in the antebellum period?}{\textbf{Religious reform, coinciding with the Second Great Awakening, stimulated the temperance movement aganist alcohol. New theories of health emerged, with many nonscientific practices like spas, dietary changes, and the study of the cranial shape to judge fitness, but also a slow advance in medical science like the realization of contagion. Education, too, was developed for children, women, the handicapped, and natives; asylums formed to treat the mentally ill and criminals, the disciplinary practices of which were theorized to apply neatly to schools. Feminist ideals developed greatly, with the Seneca Falls Convention in 1840 establishing a uniform set of women's rights and stimulating a suffrage movement; individual women made the most advancements, pushing boundaries to enter new professions.}}
        \cornell{How did religious reform establish a social order?}{\begin{itemize}
            \item Optimism from rejection of Calvinist doctrines through Unitarianism/Universalism/romanticism included Transcendentalists
            \item More significant group: Protestant revivalism
            \begin{itemize}
                \item Began w/ Second Great Awakening; diff. from Transcendentalists but shared optimism for society
                \item Charles Grandison Finney believed predestination/helplessness were counterproductive; all had personal abiliity to reach God
                \begin{itemize}
                    \item Successful in NY along Erie Canal, revival-prone "burned-over" district due to complete transformation (partly due to social transformation from canal)
                    \item Rochester, NY: religious meetings gathering women, their male relatives
                    \begin{itemize}
                        \item Most followers enjoyed economic benefits of growth but feared changes of new laborer community 
                    \end{itemize}
                \end{itemize}
            \end{itemize}
        \end{itemize}
        \textbf{Numerous groups rejected Calvinist doctrines, including the Unitarians, the Universalists, and the Romantics, but the Protestants underwent the greatest transformation. It reflected a complete rejection of beliefs of predestination, with an optimistic tone. Charles Finney spread Protestant beliefs throughout New York.}}
        \cornell{How did many fight for temperance and better health?}{\begin{itemize}
            \item Many argued that alcohol $\to$ crime, disorder, poverty; women feared burden on wives due to abusive husbands
            \item Growing alcohol supply due to $\uparrow$ grain in West and $\uparrow$ distilleries in East; isolated western areas often relied for respite
            \begin{itemize}
                \item Drank significantly more than average person today (3x) 
                \item Many alcoholics joined temperance movement
            \end{itemize}
            \item American Society for the Promotion of Temperance used revivalist techniques, relying on alcohol as a sin, to create a formal pledge against liquor
            \item Movement began to show divide in purpose
            \begin{itemize}
                \item Some banned liquor only but others wine/beer
                \item Some sought state legislature, others individual conscience to motivate 
                \item Unified in goal for moral excellence
            \end{itemize}
            \item Fight for prohibition laws generally divided Protestants and Catholics further, with alcohol generally a natural part of Catholic life
        \end{itemize}
        \textbf{Temperance began with the belief that alcohol was a sin and caused crime and poverty; as it grew in the United States, some Americans came together to form the temperance movement against alcohol. Although they were divided in specific implementations (what alcohol to ban, legal changes), they were unified in a goal for moral reform.}}
        \cornell{How did Americans explore new theories of health?}{\begin{itemize}
            \item Cholera (intestinal bacterial infection) epidemics of 1830s/1840s where less than half of those who contracted survived $\to$ health boards to find solution
            \begin{itemize}
                \item Unaware of antibiotics, modern techniques
                \item Unable to explain cause of bacterial infections
            \end{itemize}
            \item Most Americans relied on nonscientific practices
            \begin{itemize}
                \item Health spas for "water cure" (modern hydrotherapy)
                \item Sylvester Graham produced dietary theories with coarse flour (modern Graham crackers)
                \item Phrenology used study of cranial shape to determine character/intelligence
                \begin{itemize}
                    \item Analyzed indentations, bumps 
                    \item Many believed had potential to objectvely judge a person's potential
                \end{itemize}
            \end{itemize}
        \end{itemize}
        \textbf{Although new health institutions emerged to study scientific cures, far more prevalent were nonscientific theories, including health spas, dietary changes, and phrenology, or using the shape of the skull to judge skill.}}
        \cornell{How did medical science ultimately advance?}{\begin{itemize}
            \item Often appeared to lag behind rapid growth in other areas due to difficulty of experimentation, interest in poorly educated "quacks"
            \begin{itemize}
                \item Opposition to licensing in 1830s: seen as undemocratic
                \item Some traditional physicians believed in timeless truths, ancient scholars; mistrusted medical developments
            \end{itemize}
            \item Lacked basic understanding of disease 
            \begin{itemize}
                \item Edward Jenner's smallpox vaccine adapted from long-term country practice
                \item Anesthetics from New England dentist experimenting with smoother teeth removal
            \end{itemize}
            \item Significant advancement: Oliver Wendell Homes discovered contagion by analyzing spread of "puerperal fever"
        \end{itemize}
        \textbf{Although medical developments lagged behind due to physician opposition, the difficulty of experimenting on humans, the attraction of the poorly educated due to the lack of a licensing system, and numerous basic misconceptions about disease, one of the greatest medical developments in the antebellum period was the discovery of contagion.}}
        \cornell{How was education reformed in the antebellum period?}{\begin{itemize}
            \item Lacked universal public education (although states like MA had limited version), reform reflected belief of universal ability of any person to exploit innate capacity for knowledge
            \item Horace Mann, first sec. of MA Board of Education, believed educated electorate critical for democracy
            \begin{itemize}
                \item Lengthened school year, doubled teaching salaries (ignoring male/female gap), developed curriculum, introduced teacher training
            \end{itemize}
            \item Other states formed schools, teachers' colleges
            \begin{itemize}
                \item Henry Barnard in CT/RI; PA law in 1835 allowing funds for education; William Seward in NY expanded public school support 
            \end{itemize}
            \item Quality of edcation varied by state
            \begin{itemize}
                \item MA generally highly trained w/ capable educators
                \item Others lacked literate teachers or even access to education (West, Southern blacks)
            \end{itemize}
            \item Many argued for education to educate natives w/ belief in potential for civilization w/ assimilation in Far West 
            \item Educational advances numerically impressive 
            \item Mixed motivations for education (\textit{ex}: Alcott in Concord created school w/ self-taught children reflecting belief in self-realization)
            \item New institutions supported handicapped, activities known as Benevolent Empire
            \begin{itemize}
                \item Perkins School for the Blind in Boston hoped to bring inner strength to the suffering
            \end{itemize}
            \item Majority of schools aimed to instill set of social values in children
        \end{itemize}
        \textbf{Education was reformed significantly, beginning in MA under Horace Mann; other states, too, expanded their educational systems. Although many teachers were untrained and illiterate, educational developments remained numerically impressive. Furthermore, new schools emerged to aid the handicapped and instill a set of social values in children.}
        }
        \cornell{How did a growing belief in rehabilitation emerge?}{\begin{itemize}
            \item Asylums emerged for criminals/mentally ill, grouping all "criminals," notably debtors, mentally ill, paupers in disgusting conditions
            \item 1820s saw replacement of institutions w/ penitentiaries under written reform of Dorothea Dix
            \item Prisons also aimed to reform inmates, removing beliefs leading them astray
            \begin{itemize}
                \item Radicals argued that discipline could be extended to schools/factories
            \end{itemize}
            \item Asylum movement believed to help any in need, including orphans, outcast women; facilities for poor developed
        \end{itemize}
        \textbf{Initially, asylums for all criminals grouped large numbers together in abhorrent conditions; however, under the reform of Dorothea Dix and others, cleaner penitentiaries emerged designed to reform inmates; potentially, some argued, such beliefs could be extended to schools and factories; they were indeed extended to orphanages and homes for outcast women.}}
        \cornell{How did the idea of the native reservation emerge?}{\begin{itemize}
            \item U.S. policy since arrival to New World had been relocation
            \begin{itemize}
                \item Primary goal generally to remove tribes from white civilization
                \item Secondary intent often to preserve natives from white encroachment (including Andrew Jackson)
            \end{itemize}
            \item Created enclosed regions to protect, regenerate natives (much like asylums/prisons)
        \end{itemize}
        \textbf{Reservations, enclosed regions designed to protect natives, were motivated often by preservation of native society and culture, unlike relocation.}}
        \cornell{How did feminism emerge in the antebellum period?}{\begin{itemize}
            \item Women played central role in other reform $\to$ realized male-dominated society
            \item 1830s/1840s women faced sex-based restrctions with separate spheres 
            \begin{itemize}
                \item Many women resented conditions; some defied (\textit{ex}: Sarah/Angela Grimke ignored male arguments for sexual inequality); many others pushed for reform
            \end{itemize}
            \item 1840 saw significant change: American female delegates turned away by male leaders of British antislavery convention $\to$ reformers like Lucretia Mott, Stanton, Susan B. Anthony formed Seneca Falls Convention
            \begin{itemize}
                \item Compared conditions of women to slaves
                \item Created "Declaration of Sentiments" like Declaaration of Independence demanding for right to vote, complete equality 
                \item Rejected belief in separate spheres
            \end{itemize}
            \item Many feminists were Quakers due to long-term religious belief in sexual equality; Quaker-born women shocked by gender restrictions of outside world
            \item Progress somewhat limited for women as a whole, but some individual women advanced
            \begin{itemize}
                \item Elizabeth Blackwell became physician, sister-in-law Antoinette Blackwell became first woman minister; sister-in-law Lucy Stone retained maiden name
                \item Emma Willard, Catharine Beecher worked for women's education 
                \item Formed distinct form of dress known as "bloomer" to allow for free movement with little loss of modesty 
            \end{itemize}
            \item Feminists benefited greatly from associations w/ other reform movements, notably abolitionism
        \end{itemize}
        \textbf{Feminism developed with women beginning to resent their division into spheres separate from men and direct restrictions, leading to the birth of the Seneca Falls Convention, where women united to outline common rights. Individual women were able to cross sexual boundaries to push the feminist cause in slow increments.}}
        \cornell[The Crusade Against Slavery]{How did Americans begin to more directly oppose the institution of slavery?}{\textbf{The abolitionist movement, despite starting humbly with ambitions for relocation of slaves to Africa, was transformed by Garrison, demanding universal black citizenship; he was supported by countless free blacks, who not also assisted in selling subscriptions to his newspapers but also united behind powerful leaders like Frederick Douglass. Often-violent anti-abolitionists created strains within the movement, dividing them between the more moderate, gradual advocates with slow political reform and those who sought immediate emancipation (united under Garrison). Many abolitionists became increasingly radical, turning to violence and propaganda.}}
        \cornell{What were the earliest forms of opposition to slavery?}{\begin{itemize}
            \item Early opponents rarely truly inflamed or overt
            \begin{itemize}
                \item Most opposed colonization, w/ ACS (American Colonization Society) seeking to resettle African Americans in Africa/Caribbean w/o angering Southerners
                \item Sought gradual manumission with assistance in forming new slave-based societies 
                \item Some private donors, Congressional support, forming nation of Liberia for exported slaves, soon becoming independent; but ultimately negligible
            \end{itemize}
            \item Antislavery movement had begun to decline by 1830s, with number of "colonized" fewer than number born; many African Americans did not want relocation 
        \end{itemize}
        \textbf{The earliest opponents to slavery generally remained powerless, with the most poweful movement "colonization" led by the ACS, where African Americans were shipped to Africa for resettlement. However, its long-term contributions were near-negligible.}}
        \cornell{How did William Lloyd Garrison fight for abolition?}{\begin{itemize}
            \item Garrison initially assisted NJ Quaker Lundy with antislavery newspaper \textit{Genius of Universal Emancipation}
            \item Tired of Lundy's slow/gradual desire for reform $\to$ formed own newspaper, \textit{Liberator}
            \begin{itemize}
                \item Argued that slavery opponents should sympathize directly with blacks, talking not about slavery's impact on white society but instead on African society
                \item Felt proponents of colonization merely strengthened slavery by ridding country of free blacks 
                \item Believed all African Americans deserved immediate, universal citizenship 
            \end{itemize}
            \item Attracted large numbers $\to$ founded American Anti-Slavery Society in 1832, Philadelphia convention
            \begin{itemize}
                \item Formed chapters throughout nation
            \end{itemize}
        \end{itemize}
        \textbf{Garrison, despite initially advocating for gradualism, or the slow, natural emancipation, realized that slavery was to be abolished immediately with citizenship given directly to blacks. Only by sympathizing with the slaves themselves, he argued, can one become a truly powerful abolitionist. He created the American Anti-Slavery Society in 1832, which spread rapidly throughout the nation.}}
        \cornell{How did northern free blacks partake in the abolitionist movement?}{\begin{itemize}
            \item Free blacks of North generally impoverished, oppressed worse than slaves; suffered from mob violence, unable to obtain education, able to vote in only a few states; often kidnapped and returned to slavery
            \item Northern blacks consistently proud of freedom, hoping to support Southern slaves in any means possible
            \begin{itemize}
                \item Many joined Garrison in 1830s, selling \textit{Liberator} subscriptions w/in communities
            \end{itemize}
            \item Several African Americans rose up within communities to shun slavery
            \begin{itemize}
                \item David Walker created violent pamphlet shunning whites (particularly slaveowners)
                \item Sojourner Truth created religious cult in upstate NY, speaking out against slavery but promoting less violent means
                \item Frederick Douglass most notable, escaping from MD slavery to MA
                \begin{itemize}
                    \item Lectured in England, earning enough wealth to buy freedom from owner, establish antislavery newspaper \textit{North Star}
                    \item Wrote autobiography overtly shunning state of U.S..
                \end{itemize} 
            \end{itemize}
        \end{itemize}
        \textbf{Northern blacks, despite often living in conditions more oppressed than many Southern slaves, valued their liberty and supporter slaves in multiple ways. Some joined white reformers, like Garrison, to spread the message within their communities; others, however, became powerful leaders, most notably Frederick Dougalss, an escaped slave who truly mobilized the abolitionist movement within free blacks.}}
        \cornell{To what degree was anti-abolitionism a powerful social force?}{\begin{itemize}
            \item Most critics feared social turmoil brought on by abolition, with free blacks pouring into north and threatening northern stability; others feared reduction in successful trade with South
            \item Many anti-abolitionists took violent actions
            \begin{itemize}
                \item Prudence Crandall tried to open school to African American girls $\to$ arrested, humiliated with filth
                \item Mob burned down  "Temple of Liberty," abolitionist headquarters, in 1834
                \item Garrison himself captured and threatened with hanging; situation only remedied after being jailed by authorities 
                \item Elijah Lovejoy, abolitionist newspaper editor, saw presses smashed three times by angry whites
            \end{itemize}
            \item Continual resistance to abolitionism revealed that movement itself was truly passionate and transformational, resisting frequent threat and standing for beliefs despite great risk
        \end{itemize}
        \textbf{Most anti-abolitionists of the North feared the inevitable social turmoil due to the influx of free blacks into the North as well as reduced Southern productivity. Many took violent actions, humiliating, attacking, and destroying possessions of leading abolitionists. However, the movement remained strong in the face of adversity.}}
        \cornell{How did the abolitionist movement began to show internal strains?}{\begin{itemize}
            \item By the mid-1830s, violence of anti-abolitionists $\to$ many sought more moderate approaches
            \begin{itemize}
                \item Sought peaceful struggle for abolition with gradual convincing of slaveholders that instituion was fundamentally sinful 
                \item Failure to persuade $\to$ turned to govt., joined Garrisonians w/ underground railroad
                \begin{itemize}
                    \item Supreme Court w/ \textit{Prigg v. Pennsylvania} ruling that states needn't enforce law to return fugitive slaves to owners
                    \item "Personal liberty laws" forbade officials from assisting in capture of runaways 
                    \item Abolished in federal regions like D.C. and territories, but most felt Congress lacked Constitutionality
                \end{itemize}
            \end{itemize}
            \item Many others grew in radicalism, notably Garrison
            \begin{itemize}
                \item Shocked allies (like Douglass) by attacking Constitution, churches
                \item Anti-Slavery Society split up in 1840 after Garrison demanded that women be given full equality in eyes of society 
                \item Post-1840, argued for pacificism, opposition to any forms of coercion like prisons and asylums, and eventually a split between the North and the South
                \item Remained influential with unwavering morality
            \end{itemize}
            \item Never unified in political party, but inspired Liberty Party, led by Birney of KY 
            \begin{itemize}
                \item Campaigned for "free soil": no slavery in territories, but no direct abolition
                \item Ultimately attracted support in large numbers 
            \end{itemize}
        \end{itemize}
        \textbf{The abolitionist movement was divided by the mid-1840s between the more moderate ones who sought a peaceful struggle through political reform or "moral suasion," and the radicalists under Garrison, who demanded strict morals to ultimately abolish slavery. They inspired the Liberty Party, which campaigned for free soil.}}
        \cornell{How did abolitionists turn to more drastic measures to further their cause?}{\begin{itemize}
            \item Some abolitionists $\to$ more drastic measures, w/ violence by funding arms purchases, propaganda through distorted images of slavery (like in Angelica Grimke's/Theodore Dwight Weld's \textit{American Slavery as It Is: Testimony of a Thousand Witnesses})
            \item Most powerful work was fictional: Harriet Beecher Stowe's \textit{Uncle Tom's Cabin}
            \begin{itemize}
                \item Appeared first in antisalvery newspaper, but published in 1852, selling rapidly
                \item Based around tradition of sentimental novels for women, embedding message of antislavery in known form of literature
                \begin{itemize}
                    \item Expanded movement to new audience, describing victimized slaves under cruel master 
                \end{itemize}
                \item Stowe seen as northern hero
            \end{itemize}
            \item Abolitionism remained powerful despite divisions
        \end{itemize}
        \textbf{Many abolitionists turned to more drastic measures like violence and propaganda, leading uprisings by funding arms purchases and spreading both propaganda through work claimed to be nonfiction (incorrect depictions) and fiction work, notably Stowe's \textit{Uncle Tom's Cabin}.}}
    \end{document}