\documentclass[a4paper]{article}
\usepackage{tcolorbox}
\usepackage{amsmath}
\tcbuselibrary{skins}

\title{
\vspace{-3em}
\begin{tcolorbox}
\Huge\sffamily \begin{center} Chapter 30  \mbox{} \\ \huge From the "Age of Limits" to the Age of Reagan \mbox{} \\
\LARGE Finn Frankis \mbox{} \\
\Large AP US History - July 25{$^\text{th}$}, 2018 \end{center}
\end{tcolorbox}
\vspace{-3em}
}
\date{}
\author{}

\usepackage{background}
\SetBgScale{1}
\SetBgAngle{0}
\SetBgColor{red}
\SetBgContents{\rule[0em]{4pt}{\textheight}}
\SetBgHshift{-2.3cm}
\SetBgVshift{0cm}
\usepackage[margin=2cm]{geometry}

\makeatletter
\def\cornell{\@ifnextchar[{\@with}{\@without}}
\def\@with[#1]#2#3{
\begin{tcolorbox}[enhanced,colback=gray,colframe=black,fonttitle=\large\bfseries\sffamily,sidebyside=true, nobeforeafter,before=\vfil,after=\vfil,colupper=blue,sidebyside align=top, lefthand width=.3\textwidth,
opacityframe=0,opacityback=.3,opacitybacktitle=1, opacitytext=1,
segmentation style={black!55,solid,opacity=0,line width=3pt},
title=#1
]
\begin{tcolorbox}[colback=red!05,colframe=red!25,sidebyside align=top,
width=\textwidth,nobeforeafter]#2\end{tcolorbox}%
\tcblower
\sffamily
\begin{tcolorbox}[colback=blue!05,colframe=blue!10,width=\textwidth,nobeforeafter]
#3
\end{tcolorbox}
\end{tcolorbox}
}
\def\@without#1#2{
\begin{tcolorbox}[enhanced,colback=white!15,colframe=white,fonttitle=\bfseries,sidebyside=true, nobeforeafter,before=\vfil,after=\vfil,colupper=blue,sidebyside align=top, lefthand width=.3\textwidth,
opacityframe=0,opacityback=0,opacitybacktitle=0, opacitytext=1,
segmentation style={black!55,solid,opacity=0,line width=3pt}
]

\begin{tcolorbox}[colback=red!05,colframe=red!25,sidebyside align=top,
width=\textwidth,nobeforeafter]#1\end{tcolorbox}%
\tcblower
\sffamily
\begin{tcolorbox}[colback=blue!05,colframe=blue!10,width=\textwidth,nobeforeafter]
#2
\end{tcolorbox}
\end{tcolorbox}
}
\makeatother

\parindent=0pt

\begin{document}
\maketitle
\SetBgContents{\rule[0em]{4pt}{\textheight}}

\cornell[Politics and Diplomacy after Watergate]{How did American presidents following Nixon attempt to recover from the disaster of Watergate?}{}
\cornell{How did Gerald Ford attempt to recover the nation's prosperity in the aftermath of Watergate?}{
    \begin{itemize}
        \item Major setback emerged immediately due to poor decision to completely pardon Nixon
        \begin{itemize}
            \item Led many to suspect collusion between Nixon and Ford
            \item Immediate decline in popularity
        \end{itemize}
        \item Economic policies relatively unsuccessful
        \begin{itemize}
            \item Attempted to curb inflation by calling for voluntary efforts of people, rejecting idea of price/wage controls
            \item Struggled with recession intensified by energy crisis
            \begin{itemize}
                \item \textbf{Arab oil embargo} of 1973 led to extreme increase in price of oil
            \end{itemize}
        \end{itemize}
        \item Political policies often simply continuations of Nixon administration
        \begin{itemize}
            \item Signed \textbf{SALT II}, arms control accord desired by Nixon
            \item Secretary of state \textbf{Henry Kissinger} required Israel to return parts of Sinai to Egypt
        \end{itemize}
    \end{itemize}
}
\cornell{Another research question}{}
\cornell{Another research question very long for one line.}{}
\cornell{Do you want fill much more than a page}{No problem.}
\cornell{More?}{No problem.}

\cornell[Last question]{This is the end?}{Yes.}
\end{document}