\documentclass[a4paper]{article}
    \input{../notesheader.tex}
    \usepackage[normalem]{ulem}

    \newcommand{\chapternumber}{20}
    \newcommand{\chaptertitle}{The Progressives}
    \title{\vspace{-3em}
    \begin{tcolorbox}
    \Huge\sffamily \begin{center} AP US History  \\
    \LARGE Chapter \chapternumber - \chaptertitle \\
    \Large Finn Frankis \end{center} 
    \end{tcolorbox}
    \vspace{-3em}
    }
    \date{}
    \author{}
    
    \begin{document}
        \maketitle
        \SetBgContents{\rule[0em]{4pt}{\textheight}}
        \cornell[Key Concepts]{What are this chapter's key concepts?}{\begin{itemize}
            \item \textbf{6.1.I.E} - Businesses began to turn outside of the U.S. to control foreign markets/resources 
            \item \textbf{7.2.I.C} - WWI saw restrictions on freedom of speech w/ fears of radicalism $\to$ attacks on immigrants, unions
            \item \textbf{7.2.II.B} - $\uparrow$ demand for production/labor during WWI/WWII, Great Depression $\to$ $\uparrow$ urban migrations 
            \item \textbf{7.2.II.C} - Great Migration during/after WWI saw Afr. Americans departing South for North/West to escape discrimination; still far from perfect
            \item \textbf{7.3.II.A} - Initial U.S. WWI neutrality reversed by Wilson to defend democratic principles of nation 
            \item \textbf{7.3.II.B} - American Expeditionary Forces never played major combat role; arrival of U.S. $\to$ conflict began to favor Allies
            \item \textbf{7.3.II.C} - U.S. Senate would not ratify Treaty of Versailles or join League of Nations despite Wilson's signif. involvement in their creation
        \end{itemize}}
        \cornell[America and the World: 1901-1917]{How did America interact with the foreign world before WWI?}{\textbf{Roosevelt focused on the distinction between civilized and uncivilized nations as well as attempting to limit disputes between other nations; however, he was never afraid to show U.S. power as a threat. In Latin America, Roosevelt established the U.S. right to intervene in unstable Latin American governments; he also ensured the construction of the Panama Canal by encouraging a Panamanian Revolution. Taft focused on strengthening American investments in outside territories. Wilson hoped to take a more moral stance on diplomacy; however, he was never afraid to establish a military government or go to war with another nation over a conflict of ideology.}}
        \cornell{How did Roosevelt argue the importance of civilized nations?}{\begin{itemize}
            \item Roosevelt felt "civilized" nations defined by race (white-dominated) as well as economic proninence (\textit{ex}: Japan would be considered "civilized" for its great prosperity)
            \item Natural duty of civilized societies to assist/intervene in "backward" nations $\to$ inspired rapid growth of American navy
        \end{itemize}
        \textbf{Roosevelt used racial and economic justifications for the natural superiority of certain "civilized" nations. It was the fundamental duty of such nations to assist the more "backward" ones.}}
        \cornell{How did Roosevelt protect Asian free trade?}{\begin{itemize}
            \item Russo-Japanese War stimulated by surprise Japanese attack mediated by Roosevelt at request of Japanese
            \begin{itemize}
                \item Roosevelt formalized Japanese territorial gains as well as received Japanese agreement to cease aggression
                \item Secretly agreed to preserve free trade w/ Japanese
            \end{itemize}
            \item Roosevelt respected for work as mediator (Nobel Peace Prize); relations quickly deteriorated w/ Japan becoming dominant Pacific naval power $\to$ pushed out U.S. from trading w/ territories 
            \begin{itemize}
                \item U.S. asserted naval dominance w/ show of "Great White Fleet," demonstrating immense naval power
            \end{itemize}
        \end{itemize}
        \textbf{Roosevelt helped mediate the Russo-Japanese War, benefiting the U.S. by receiving a secret free trade agreement with the Japanese. As Japan grew more prosperous, they began to exclude the U.S. from several ports, forcing the U.S. to assert their naval dominance to Japan with a bold display of power.}}
        \cornell{How did the U.S. interact with Latin America?}{\begin{itemize}
            \item 1902: Venezuelan govt. unable to pay back debts to Euro. bankers $\to$ GB, Italy, Germany blockaded coast w/ Germany bombarding port $\to$ U.S. naval threat pushed away
            \item 1904: Roosevelt established \textbf{Roosevelt Corollary} to Monroe Doctrine stressing ability of U.S. to intervene in \underline{unstable governments} of Western Hemisphere
            \begin{itemize}
                \item Motivated by Dominican Republic crisis, unable to pay back debts to Europe after internal revolution $\to$ U.S. controlled customs, keeping 45\% of revenue internal, 55\% to Euro. creditors 
                \item Cuba's agreement to Platt Amendment (U.S. could intervene in foreign affairs) saw U.S. grant independence; U.S. troops quickly intervened in 1906 w/ domestic uprisings
            \end{itemize}
        \end{itemize}
        \textbf{Roosevelt passed a corollary to the Monroe Doctrine which allowed the U.S. to intervene in unstable governments of the Western Hemisphere. It justified American intervention in Dominican Republic customs as the government faced bankruptcy as well as a domestic Cuban uprising.}}
        \cornell{How did Roosevelt oversee the construction of the Panama Canal?}{\begin{itemize}
            \item Roosevelt turned to Isthmus of Panama in Colombia for canal linking Atlantic and Pacific due to short distance (despite not being at sea level $\to$ extra cost for locks), existing (but failed) construction by French company 
            \item Sent John Hay (sec. of state) to negotiate w/ Colombian govt., pressuring into agreement to give U.S. six-mile "canal zone"
            \begin{itemize}
                \item Colombian senate refused, demanding higher price
                \item Roosevelt encouraged revolution in Panama w/ new govt. independent of Colombia; resisted Colombian attempts at putting down 
                \item Panamanian govt. agreed to terms w/ work soon beginning
            \end{itemize}
            \item Panama Canal opened in 1914
        \end{itemize}
        \textbf{Roosevelt sought to link the Atlantic and Pacific with a sea route; he turned to the Colombian Isthmus of Panama. After the Colombian senate failed to agree to U.S. terms, Roosevelt encouraged a revolution in Panama, creating a new independent nation which quickly agreed to the desired terms.}}
        \cornell{How did Taft employ his "Dollar Diplomacy" policy?}{\begin{itemize}
            \item Taft had little interest in Roosevelt's broad goal for world stability; instead sought to bring US investments to other regions
            \item Targeted Caribbean: after 1909 rev. in Nicaragua, sided w/ rebels, sending troops to restore peace
            \begin{itemize}
                \item American bankers offered loans to new Nicaraguan govt. $\to$ financial power
                \item U.S. troops sent in and remained to protect govt.
            \end{itemize}
        \end{itemize}
        \textbf{Taft's "Dollar Diplomacy" policy sought to promote U.S. investment interests worldwide, particularly in the Caribbean, where he supported and offered loans to Nicaraguan insurgents.}}
        \cornell{How did Wilson embody the principle of "Moral Diplomacy"?}{\begin{itemize}
            \item Wilson faced large international challenges despite little experience in foreign affairs; mostly strengthened Roosevelt-Taft policies
            \item Wilson established mil. govt. in Dom. Repub.; sent marines to calm rev. in Haiti, remaining until 1934
            \item Bought Danish West Indies after fear that Germany would take over; signed treaty w/ Nicaragua agreeing to bar all other nations from building canals there
            \item In Mexico (w/ large American business presence), leader Díaz had been overthrown by popular Madero; U.S. govt. encouraged Huerta to depose him, w/ Taft admin. preparing to recognize Huerta govt.
            \begin{itemize}
                \item Huerta murdered Madero shortly before Wilson $\to$ refused to recognize as valid govt.
                \item Wilson hoped refusing recognition would end regime; American businessmen agreed w/ mil. regime $\to$ using pretense of Mexican army incorrectly arresting U.S. soldiers, giving insufficient apology to seize port of Veracruz 
                \item Wilson's attack on Veracruz $\to$ Carranza (Constitutionalist) took control, but Wilson remained unsatisfied due to limited support of U.S. guidelines
                \item Forced to recognize Carranza after orig. U.S. ally, Pancho Villa, greatly weakened $\to$ Villa felt betrayed, shooting U.S. miners in Mexico, killing several more past NM border 
                \item U.S. troops unsuccessfully pursued Carranza, engaging only in a few spars w/ Carranza's army $\to$ war betw. U.S. and MX seemed close until Wilson withdrew due to world war
            \end{itemize}
        \end{itemize}
        \textbf{Wilson took an active role in foreign affairs, never afraid to send the military to a potentially dangerous or economically beneficial region, including the Dominican Republic, Haiti, and the Danish West Indies. In Mexico, constant conflict unfolded as governments rapidly transitioned; after the U.S. abandoned an early Mexican ally and they pursued him for his retaliation, the U.S. and Mexico seemed once again close to war.}}
        \cornell[The Road to War]{How did America initially approach the Great War?}{\textbf{War broke out in 1914 in Europe due to the long-term rivarly between Britain and Germany. Wilson initially proclaimed American neutrality, but struggled greatly due to British economic ties, angering the Germans and promoting attacks on U.S. vessels; he even won the reelection by emphasizing his policy of neutrality. However, angered by continual German aggression, he ultimately chose to prepare for the war in 1915 and entered the war on April 2nd, 1917.}}
        \cornell{How did war break out in Europe?}{\begin{itemize}
            \item By 1914, two alliances: "Triple Entente" of GB, FR, RU and "Triple Alliance" of Germany, Austria-Hungary, Italy
            \begin{itemize}
                \item Main rivalry betw. GB/Germany as Germany began to approach Britain in colonial and naval supremacy 
            \end{itemize}
            \item Immediate conflict: Austro-Hungarian Archduke Franz Ferdinand assassinated in Sarajevo in June 1914 by Serbian nationalist
            \begin{itemize}
                \item Germany supported Austria-Hungary in attacking Serbia; Serbia requested Russian assistance
                \item Germany had declared war on Russia/France by August $\to$ invaded Belgium; GB quickly joined war to honor alliance w/ France, attack Germany 
                \item Italy initially neutral, later w/ Triple Entente; Ottoman Empire joined later $\to$ entire continent, part of Asia engaged in war 
            \end{itemize}
        \end{itemize}
        \textbf{The long-term cause of the war was a rivalry between Britain and Germany due to their competing colonial interests. The war was immediately stimulated by the assassination of Franz Fedinand; nations began to take sides and soon became embroiled in war.}}
        \cornell{How did Wilson initially proclaim neutrality in the war?}{\begin{itemize}
            \item Wilson initially declared neutrality, but supported British cause (along w/ many others) due to admiration of GB, skilled propaganda; some German/Irish Americans supported Germany
            \item Strong econ. ties to Britain, unable to trade w/ Germany due to British naval blockade; could afford to cut ties w/ Germany, other Central Powers but not w/ Britain and France (particularly due to $\uparrow$ wartime demands)
            \item Germans began to attack vessels shipping goods to GB using submarines; after sinking British passenger ship carrying 128 Americans, great anger; Wilson demanded German commitment to neutrality w/ agreement but very reluctant
            \item After Allies began to sink German submarines, Germany began attacking unarmed French boats $\to$ Americans injured
        \end{itemize}
        \textbf{Although the U.S. declared neutrality, their trading ties with Britain made it increasingly difficult. Seeking to cut off supplies, Germany often attacked American shipping vessels, once even sinking a British passenger vessel and killing 128 Americans. Tensions began to rise between the two nations.}}
        \cornell{How did Wilson partake in the debate between preparedness and pacifism?}{\begin{itemize}
            \item Facing reelection, Wilson was forced to address question of war; initially sided w/ antipreparedness forces but later changed mind w/ fall 1915 seeing increase in armed forces, working hard for Congressional approval
            \item Congressional peace side saw $\uparrow$ strength $\to$ Wilson began to argue that his work had kept nation out of war; opponent, Charles E. Hughes, far more likely
            \item Wilson won election narrowly
        \end{itemize}
        \textbf{Wilson eventually decided it was best for the nation to prepare for the potential of war, greatly expanding the size of the American navy despite Congressional opposition. In winning reelection, he stressed that his work had helped keep the nation safe and away from war.}}
        \cornell{How did the war ultimately begin as a battle for democracy?}{\begin{itemize}
            \item Wilson took idealistic approach: felt U.S. had no material benefit, rather fighting for new world order for peaceful league of nations after conclusion of war
            \item Several provocations essentially forced intervention
            \begin{itemize}
                \item Germans sought to sink all Allied defense ships before U.S. could intervene
                \item Foreign minister Zimmermann sent telegram to Mexico asking to join war if U.S. joined, in exchange for all "lost provinces" in Southwest
                \item Overthrow of Russian monarchy for republican govt. $\to$ no shame in intervening
            \end{itemize}
            \item Wilson gave powerful speech to Congress asking to join war; some opposition but eventually approved
        \end{itemize}
        \textbf{Wilson joined the war for the sake of promoting democratic ideals: American entry to war was promoted by the continual German attacking of U.S, and Allied ships, the German offer to Mexico in the Zimmermann Telegram for Mexico to join the war against the U.S. in exchange for their lost territories, and the democratization of Russia.}}
        \cornell["War Without Stint"]{What characterized American intervention in the war?}{\textbf{The U.S. initially entered the war hoping to end things with naval assistance; the Russian surrender meant that ground troops were required. Wilson recruited ground troops primarily through forced conscription; women and African Americans were permitted to join the army, too. The U.S. army slightly tipped the tide, allowing the war to end within eight monhts of their arrival. The technology of the war stimulated great devastation.}}
        \cornell{How did the U.S. enter the war?}{\begin{itemize}
            \item German submarines $\to$ GB lost $\approx \dfrac{1}{4}$ of supplies $\to$ U.S. helped attack German submarines (through direct attack, mines), escort merchant vessels w/ warships
            \begin{itemize}
                \item Dramatic reduction in loss of Allied goods
            \end{itemize}
            \item With Nov. 1917 Bolshevik Revolution in Russia $\to$ neutrality w/ Central Powers, Americans forced to send ground troops to assist weakened Allies
        \end{itemize}
        \textbf{Initially, the Americans helped primarily at sea, attacking German submarines and escorting merchant vessels with warships. However, after Russia negotiated neutrality with Germany, America was forced to send ground troops.}}
        \cornell{How did America prepare to mobilize ground troops?}{\begin{itemize}
            \item 1917 saw 120k soldiers in army, 80k in National Guard; neither had combat experience, very few had battle experience
            \item Several sought voluntary recruitment (including ailed Roosevelt); Wilson decided on forced conscription w/ Selective Service Act in May: brought 3m men into army and 2m men joining voluntarily
            \begin{itemize}
                \item Made up American Expeditionary Force (AEF)
            \end{itemize}
            \item Represented first time in U.S. history w/ signif. overseas battle; morale difficult to boost in polluted trenches
            \begin{itemize}
                \item Most soldiers went to bars/brothels in new towns $\to$ venereal disease widespread
            \end{itemize}
            \item Women permitted to enlist in AEF (not combat) in hospitals/offices
            \item Nearly 400k Afr. Americans enlisted/drafted in army/navy (barred from marines); 50k went to Europe in segregated, white-commanded troops
            \begin{itemize}
                \item Most remained at home, performing menial tasks $\to$ some revolts against oppressive whites
            \end{itemize}
            \item Widespread "IQ" testing for soldiers; measured education, not intelligence $\to$ half of whites, most of Afr. Americans scored very low
        \end{itemize}
        \textbf{Wilson dismissed voluntary conscription, instead encouraging a draft to obtain a reasonable number of fighters. The overseas battle saw great lowering in morale, only boosted by local bars and brothels. Both women (no combat) and African Americans enlisted in the army; African Americans were generally forced to perform menial tasks.}}
        \cornell{What characterized American fighting in the Great War?}{\begin{itemize}
            \item Signif. numbers of U.S. troops arrived by spring of 1918; war ended eight months later
            \item Army led by Pershing (commander in unsuccessful pursuit of Pancho Villa); command structure independent of Allies
            \item U.S. troops surrounded by others who had been fighting for nearly four years in violent, uncomfortable, boring, fearful trenches $\to$ limited morale 
            \item U.S. troops had very brief time in trenches: turned the table to allow allies to break out of trenches, advance 
            \begin{itemize}
                \item June 1918: U.S. forces at Château-Thierry assisted French, again at Rheims; offensive began by July 18th
                \item Sept. 1918: joined assault against Germans in Argonne Forest, pushing back nearly to border after seven weeks
            \end{itemize}
            \item Germany risked invasion $\to$ sought armistice (cease-fire) to stimulate negotiations; U.S. wanted to enter Germany but exhausted Allied leaders agreed $\to$ war ended in Nov. 1918
        \end{itemize}
        \textbf{When the U.S. ground troops arrived, morale was already greatly weakened among the Allied troops due to the crushing devastation characterized by trench warfare. However, the U.S. quickly turned the tide, advancing with the Allies on the Germans; as the troops approached the German border, Germany sought an armistice and the war ended.}}
        \cornell{How was the Great War characterized by new, devastating technology?}{\begin{itemize}
            \item Trench war necessary due to devastating modern warfare like machine guns $\to$ open warfare too deadly
            \begin{itemize}
                \item Led to creation of chemical weapons to attack trenches $\to$ troops constantly carried gas masks
                \item Technological warfare required far more maintenance w/ more ammunition for machine guns, fuel for vehicles $\to$ often difficult for forces to arrange rapid delivery of equipment 
            \end{itemize}
            \item WWI first airplane-based conflict: simple planes still effective due to limited anti-aircraft tech. $\to$ mainly for bombing, fighting, reconnaissance
            \item Navy most modernized: GB \textit{Dreadnought} characterized by hydraulic guns, electric power, wireless telegraph, navigation; submarines driven by powerful, long-range diesel engine (notably German U-boats) very significant
            \item Technology caused high number of casualties, with several million lost from each major nation; U.S. only lost 112k but half due to flu; battles w/ critical U.S. troops saw high casualties  
        \end{itemize}
        \textbf{The devastating technology of the Great War required trench warfare rather than the devastating open warfare; though airplanes became significant for the first time in a war, the navy was the most modernized, with powerful submarines and vessels. Technology was the cause of the high casualties.}}
        \end{document}